---
title: "Assessment for SRM"
description: |
  More details on your assessment
author:
  - name: Ralf Martin
    url: https://mondpanther.github.io/wwwmondpanther/
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    self_contained: true
    css: webex.css
    includes:
      after_body: webex.js
editor_options: 
  chunk_output_type: inline
---





```{r setup, include = FALSE, message=FALSE}
library("webexercises")
library(ggplot2)
#knitr::opts_chunk$set(fig.path='../../images/')
```





# Exam


- 2 Open (Essay style) Answer Questions: select 1 ,  20 Marks

- 32 Multiple Choice Questions: 4 answer options, 2.5 Marks each, with no deduction for a wrong answer.

### Example for open questions

- see [Question 10.4](https://mondpanther.github.io/datastorieshub/posts/exercises/exercises10/#pollution) Question from Exercise sheet 10



### Multiple Choice Question Examples
(To check your understanding select what you think is the correct option in the dropdown selection menus. If you are right the menu box should turn blue)

#### Question 1
Suppose you have estimated the following regression model on a representative sample of the UK population
$$\log(Weight_i)=\beta_0+\beta_1 Female_i+\beta_2 AGE_i +\beta_3+AGE_i^2+\epsilon_i$$

where Weight_i is weight of a person in 100s of kilograms
Suppose you find that $\hat\beta_1=-0.10$ Which of the following is a correct interpretation?

(a) The average weight for women is 10kg
(b) Women – on average - weigh 10 kg less than men
(c) Women – on average – weigh 10% less than men
(d) None of the above.
	
Correct answer `r mcq(c("a","b",answer="c","d"))`

<br>

#### Question 2

For the model from Question 1: suppose you find that $\hat\beta_2=0.01$ and $\hat\beta_3=-0.0001$ .
What does it tell you about the age at which we expect people to be heaviest?

(a) At 25 years
(b) At 75 years 
(c) At 40 years
(d) None of the above

Correct answer `r mcq(c("a","b","c",answer="d"))`

`r hide("Hint")`
Because $\beta_3$ is negative the weight-age relationship is hump shaped. We can find the
maximum of the hump by setting the gradient equal to 0 implying

$$0.01-2\times 0.0001\times AGE=0$$
Consequently:

$$AGE=\frac{0.01}{2\times 0.0001}=50$$


`r unhide()`

<br>

#### Question 3

Using a univariate model you obtain a parameter estimate $\hat\beta=2.1$ with a standard error of 0.1 obtained using a sample size of 1000 observations. Should you reject the hypothesis that β=1 at the 5% level?

(a)	Yes
(b)	No
(c)	It depends on the variance of the residuals
(d)	It depends on the variance of the explanatory variable

`r mcq(c(answer="a","b","c","d"))`

`r hide("Hint")`
The t value becomes $t=\frac{2.1}{0.1}=$`r 2.1/.1` , which is larger than 1.96, the 5% critical threshold
with a standard normal distribution. Note that 1000 observations we can ignore that the t-statistic is only approximately normally distributed.

`r unhide()`


<br>

#### Question 4

The demand for a new drug is known to be linear and downward sloping; i.e. a higher price means a lower demand. A researcher provided an estimate of this demand curve but suspects that a confounding factor led to a downward bias. This means that the estimated curve is

(a) Flatter than it should be
(b)	Steeper than it should be
(c)	Upward sloping
(d)	None of the above

`r mcq(c("a",answer="b","c","d"))` 

<br>

#### Question 5

The probability that a standardised normal variable takes a value of zero or more is: 

(a)	Equal to zero.
(b)	50%
(c)	95% for a significance level of 0.1
(d)	I need more information on the variance and probability distribution to answer this question. 

`r mcq(c("a",answer="b","c","d"))` 


#### Question 6
```{r ,include=FALSE,message=FALSE}

#clean and combine the dataset
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(car)
library(sandwich)
library(AER)
rank<-read.csv("https://raw.githubusercontent.com/zihanz14/data_team17/main/ForbesAmericasTopColleges2019.csv")
student<-read.csv("https://raw.githubusercontent.com/zihanz14/data_team17/main/studentsss.csv")
data1<-inner_join(rank,student)
case<-read.csv("https://raw.githubusercontent.com/zihanz14/data_team17/main/colleges.csv")
data2<-inner_join(data1,case)
datafinal<-data.frame(college=data2$college,rank=data2$Rank,outstaterate=data2$Percent.out.of.state/100,student=data2$Undergraduate.Population,case=data2$cases,county=data2$county)
mask<-read.csv("https://raw.githubusercontent.com/zihanz14/data_team17/main/mask.csv")
data3<-mask %>% group_by(county) %>% summarise(invalidmask=mean(rate.of.people.who.use.masks.uneffectively))
datafinal1<-inner_join(datafinal,data3)
datafinal1 <- datafinal1[!is.na(datafinal1$case),]
party<-read.csv("https://raw.githubusercontent.com/zihanz14/data_team17/main/party.csv")
dataparty<-inner_join(party,case)


datafinal2<-inner_join(party,datafinal1)
datafinal2 <- datafinal2[!is.na(datafinal2$case),]
datafinal2=datafinal2 %>% mutate(casesOstudent=case/student)
```


The R output below shows a regression of COVID19 cases per student among US Universities in Fall 2020. The variable `partyrank`
ranks Universities according to the quality of the local party scene (i.e. the uni with the best partyscene has rank 1).
What does the regression suggest abou the relationship between party rank and covid cases?


```{r}
#simple regression of case on partyrank
lm(casesOstudent~partyrank,datafinal2) %>% summary()


```
Moving to one lower rank (e.g from rank 4 to rank 5) ...

(a)	... leads to 5.2 students less being affected by covid
(b)	... leads to 5.2% less COVID cases.
(c) ... leads to 0.52 less COVID cases in 10,000 students.
(d)	... leads to 5.2 less COVID cases in 1000 students.

`r mcq(c("a","b",answer="c","d"))` ..




#### Question 7

The figure below shows the result of a Monte-Carlo study of a parameter estimate. It's a density plot of the estimated parameter for a large number of replications along with the true parameter value (vertical line)


```{r,include=TRUE}
obs=1000
df=data.frame(mc=rnorm(obs)+1)
library(ggplot2)
ggplot(df,aes(x=mc))+geom_density()+geom_vline(xintercept=2,color="red")+theme_minimal()+xlab("")

```
(a)	The estimate is unbiased.
(b)	The estimate is upward biased.
(c) The estimate is downward biased
(d)	There is not enough information to tell

`r mcq(c("a","b",answer="c","d"))` 


#### Question 8

```{r}
library(readxl)
set <- read_excel("datastories/worldoilprice/Dataset.xlsx",  skip = 0) %>% 
       rename(oilprice=`WTIspotprice`)
```

Consider the R output below where the variable `oilprice` is the spot price for oil computed at the monthly level from 2010-19. Based on what is reported below, which of the following can you conclude:

(a)	The oil price is non-stationary at a 5% significance level but not at 1%.
(b)	The oil price is stationary at all significance levels.
(c) The oil price is stationary only at the 10% significance level.
(d) The oil price non-stationary at all significance levels.

`r mcq(c("a","b","c",answer="d"))` 


```{r,include=TRUE}
library(urca)
ur.df(set$oilprice) %>% summary()

```


#### Question 9

Consider the R output below where the variable `oilprice` is (as in Question 8) the spot price for oil computed at the monthly level from 2010-19. Based on what is reported below, which of the following can you conclude:

(a)	Oil prices are growing at a rate of 7% on average.
(b)	Oil prices are shrinking at a rate of 3.1% on average.
(c) Oil prices are shrinking at a rate of 0.31% on average.
(d) Oil prices are srhinking at a rate of 0.23% on average.

`r mcq(c("a","b","c",answer="d"))` 

```{r}

lm(formula=diff(log(oilprice))~1,set) %>% summary()


```
#### Question 10

Which notation is correct? 

(a) $$y_i=β_1+β_2 x_{2i}$$   
(b) $$\hat{y}_i=\hat{β}_1+\hat{β}_2 x_{2i}$$   
(c) $$\hat{y}_i=β_1+β_2 x_{2i}$$
(d) $$\hat{y}_i=\hat{β}_1+\hat{β}_2 \hat{x}_{2i}$$   


`r mcq(c("a",answer="b","c","d"))` 
