---
title: "Quick Guide - Log linear models"
author: Ralf Martin
output:
  distill::distill_article:
    self_contained: false
    css: ../../webex.css
    includes:
      after_body: ../../webex.js
categories:
  - Quickguides
editor_options: 
  chunk_output_type: inline
---


```{r setup, include = FALSE, message=FALSE}
library("webex")
library(ggplot2)
#library(tutorial)
#tutorial::go_interactive()
#go_interactive( greedy = FALSE, height = 200 )

```

What is the difference between the linear, log and log-log models? They describe different possible relationships between the dependent and explanatory variable. This affects how we can interpret estimated parameters. There are three cases

- Linear: $y=a+bX$
- Log:  $\ln{y}=a+bX$
- Log-log:  $\ln{y}=a+b \ln{X}$


We can draw all three in the same y-x diagram by re-writing the equations (and recalling that the exponential function $\exp(\cdot)$ is the inverse of the natural logarithm function $\ln(\cdot)$


- Linear: $y=a+bX$
- Log:  $y=\exp(a+bX)$
- Log-log:  $y=\exp(a+b\ln{X})$

Plotting this for a=1   and b=0.1 will look as follows:

```{r,echo=FALSE, message=FALSE}
a=1
b=0.1
library(ggplot2)
p <- ggplot(data = data.frame(x = 0), mapping = aes(x = x))
fun.1 <- function(x) a+b*x
fun.2 <- function(x) y=exp(a+b*x)
fun.3 <- function(x) y=exp(a+b*log(x))
p+ stat_function(fun = fun.1,n = 1000,aes(color="Linear")) +
   stat_function(fun = fun.2,n = 1000,aes(color="Log")) +
   stat_function(fun = fun.3,n = 1000,aes(color="Log-Log")) + 
   xlim(-1,10)+ylim(0,8) + theme_minimal() +
   scale_colour_manual("Legend", values = c("red", "blue", "green", "orange"))
```


i.e. log vs log-log are two different types of a non-linear relationship between x and y; e.g. in the log case a given change in x will have an increasingly bigger impact on y. The opposite is true in the log-log case. 
The interpretation of the parameters also changes in the three cases. To understand the differences consider the derivative (or gradient) in each case:


- Linear: $\frac{\partial y}{\partial x} = b$
- Log: $\frac{\partial y}{\partial x} = \exp(a+b x)b=y\times b$
- Log-Log: $\frac{\partial y}{\partial x} = \exp(a+b \ln x)b=y\times b \times \frac{1}{x}$

For this we use the following rules for taking derivatives:

- Derivative of $\ln z$ is $\frac{1}{z}$
- Derivative of $exp(z)$ is $exp(z)$ 
- Chain rule: $\frac{\partial(f(g(x)))}{\partial x} = \frac{\partial(f(g(x)))}{\partial g(x)} \times \frac{\partial g(x)}{\partial x}$


To understand how to interpret the b coefficient, we can re-write the derivative formulas so that we have b on one side of the equation. We can also use the fact that the derivative is approximately the small change in y you get in response to a small change in x; i.e. $\frac{\partial y}{\partial x} \approx \frac{dy}{dx}$

- Linear: $b=\frac{\partial y}{\partial x} \approx \frac{dy}{dx}$
- Log: $b=\frac{\partial y}{\partial x} \frac{1}{y}\approx \frac{dy}{y} \times \frac{1}{dx}$
- Log-log: $b=\frac{\partial y}{\partial x} \frac{x}{y}\approx \frac{\frac{dy}{y}}{ \frac{dx}{x}}$


Hence, we get the following interpretations (recalling that $dz/z$ is the growth rate of a variable z):

- Linear:	Change in $y$ if we change $x$ by 1 unit
- Log:	Growth rate of $y$ if we change $x$ by 1 unit (the growth rate in percent will be b times 100)
- Log-log:	Growth rate of $y$ (in percent) in response to growth of $x$ by 1 percent. Note that if $x$ grows by 1 percent we have that $dx/x=0.01$

### Examples:

Suppose that y is the wage in \$ and  $x$ is the years of experience of an individual

- Linear:	1 year more of experience means b$ more wage
- Log:	1 year more of experience means b×100 percent more wage
- Log-log:	If your experience goes up by 1% (e.g. if you current experience is 10 years, 1% is 0.1 years or about 1 months 1 one week) then your wage will go up by b%.

Suppose that y is the number of crimes and is the number of foreigners in an area

- Linear:	1 foreigner more implies b more crimes
- Log:	1 foreigner more implies $b\times 100$ percent more crimes
- Log-log:	If the number of foreigners goes up by 1% then the number of crimes goes up by b%.





<!--
```{r}
knitr::include_url("https://mondpanther.github.io/datastorieshub/posts/quickguides/quickguide_loglinear/quickguide_loglinearold.htm")
```
-->
