---
title: "Quick Guide - Glossary of R commands"
author: Ralf Martin
output:
  distill::distill_article:
    self_contained: false
    css: ../../webex.css
    includes:
      after_body: ../../webex.js
categories:
  - Quickguides
editor_options: 
  chunk_output_type: inline
---


```{r setup, include = FALSE, message=FALSE}
library("webex")
library(ggplot2)
#library(tutorial)
#tutorial::go_interactive()
#go_interactive( greedy = FALSE, height = 200 )

```




<!--html_preserve-->
<script src="https://cdn.datacamp.com/datacamp-light-latest.min.js"></script>

<style>
  /* Rearrange console label */
  .datacamp-exercise ol li, .datacamp-exercise ul li {
    margin-bottom: 0em !important;
  }
  
  /* Remove bullet marker */
  .datacamp-exercise ol li::before, .datacamp-exercise ul li::before {
    content: '' !important;
  }
</style>
<!--/html_preserve-->




## `for()`
`r hide("More...")`
Repeating a series of commands several times

#### Example
<!--html_preserve-->
<div data-datacamp-exercise data-lang="r" data-height="150">
  <code data-type="sample-code">
  for(ii in 1:10){
     cat("Round")
     cat(ii)
     cat("\n")
  }
  </code>
</div>
<!--/html_preserve-->

[More Information](https://www.datamentor.io/r-programming/for-loop/)

`r unhide()`




## `library()`

`r hide("More...")`

Loads an extension package into memory.

#### Example

```{r,echo=TRUE}
library(car)

```

Loads the car library that allows you to run the `linearHypothesis()` command (and more).

[More Information](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/library)

`r unhide()`


## `lm()`

`r hide("More...")`

Implements the a linear regression model using the least squares algorithm


#### Example


<!--html_preserve-->
<div data-datacamp-exercise data-lang="r" data-height="150">
  <code data-type="sample-code">
    library(AER)
    library(dplyr)
    data("Affairs")  # Loads `Affairs` dataframe into memory (part of AER library)
    
    reg=lm(affairs~age+gender,data=Affairs)
    
    reg %>% summary()
  
  </code>
</div>
<!--/html_preserve-->
`r unhide()`



## `read.csv()`

`r hide("More...")`

Reads csv files (comma separated value; i.e. a basic table format) from your harddrive or the web into an R dataframe.


#### Example

```{r, echo=TRUE}
df=read.csv("https://www.dropbox.com/s/a2opu10e2hz0dps/brexit.csv?dl=1")

```

Loads the `brexit.csv` dataset


`r unhide()`

