---
title: "Quick Guide - Glossary of R commands"
author: Ralf Martin
output:
  html_document:
    self_contained: false
    css: ../../webex.css
---

```{r setup, include = FALSE, message=FALSE}
#library("webex")
library(ggplot2)
library(tutorial)

#tutorial::go_interactive()


```




```{r,eval=FALSE, echo=FALSE, results='asis', purl=FALSE}
cat('
<script src="https://cdn.datacamp.com/datacamp-light-1.0.0.min.js"></script>
')
 
```

<script src="https://cdn.datacamp.com/datacamp-light-1.0.0.min.js"></script>



<div data-datacamp-exercise data-lang="r">
    <code data-type="pre-exercise-code">
        # This will get executed each time the exercise gets initialized
        b = 6
    </code>
    <code data-type="sample-code">
        # Create a variable a, equal to 5


        # Print out a


    </code>
    <code data-type="solution">
        # Create a variable a, equal to 5
        a <- 5

        # Print out a
        print(a)
    </code>
    <code data-type="sct">
        test_object("a")
        test_function("print")
        success_msg("Great job!")
    </code>
    <div data-type="hint">Use the assignment operator (<code><-</code>) to create the variable <code>a</code>.</div>
</div>

```{r, echo=FALSE, results='asis', purl=FALSE}
cat('


')
 
```




## `library()`

Loads an extension package into memory.

#### Example

```{r}
library(car)

```

Loads the car library that allows you to run the `linearHypothesis()` command (and more).

[More Information](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/library)



## `lm()`

Implements the a linear regression model using the least squares algorithm


#### Example

```{r}
library(AER)
library(dplyr)
data("Affairs")

reg=lm(affairs~age+gender,data=Affairs)

reg %>% summary()

```
Loads the `Affairs` dataset from the AER library to run a regression of the number of affairs on `age` and `gender`


## `read.csv()`

Reads csv files (comma separated value; i.e. a basic table format) from your harddrive or the web into an R dataframe.


#### Example

```{r, echo=TRUE}
df=read.csv("https://www.dropbox.com/s/a2opu10e2hz0dps/brexit.csv?dl=1")

```

Loads the `brexit.csv` dataset




