---
title: "Exercises 1"
description: |
  Starting to think like an econometrician.
author:
  - name: Ralf Martin
    url: https://mondpanther.github.io/wwwmondpanther/
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    self_contained: true
    css: ../../webex.css
    includes:
      after_body: ../../webex.js
categories:
  - Exercises
editor_options: 
  chunk_output_type: inline
---



```{r setup, include = FALSE, message=FALSE}
library("webex")
library(ggplot2)
knitr::opts_chunk$set(fig.path='../../images/')
```



In the lecture we discussed the notion of causality. We contrasted the causality mechanism that you might be most interested in (your story) with the causality mechanism(s) that might be driving your data (the data story). For good data-analysis you need to be aware of both: Have a clear idea of what you want to analyse or explore as well as of all the things that could be going on in your data. To practice this I ask you to go through a couple of examples.
These are suggestive causal relationships typically found on the basis of some correlation  in in one dataset or another. 
In each case:

- Can you think of a mechanism that would support the causal relationship suggested? 
- Can you also think of at least one reason of why the relationship might have been the consequence of confounding drivers in the data or indeed reverse causality? As consequence the effect suggested might be either non-existent or biased?
- Can you also discuss the direction of the bias that would be introduced by your suggestion?

For each case I give you a suggestion as example when clicking the answer tab. You might also come across some interesting causal claims that are made e.g. in the news. If so you are very welcome to share those on the course forum.

<!-- These are typically strong correlations that have been found in one dataset or another. For each example, could you think of a mechanism or reason that would support the causal chain to work in both directions: firstly the one explicitly  You might want to share some of your ideas on the course forum. For each case I give you a suggestions as example when clicking the answer tab.
--->

# Exercise 1.1

Police body worn cameras increase violence against the police?

```{r}
knitr::include_url("https://www.govtech.com/public-safety/Study-Finds-Negative-Effects-of-Police-Worn-Body-Cameras.html")
```


`r hide("Example Answer:")`

We would typically expect the opposite; i.e. if people are on camera it is more likely that they can be prosecuted and punished for attacking a police officer. But as I suggested: a key skill of an applied data analyst is to think of imaginative stories. So here we can test that skill: Maybe there are people who like a reputation as misfits. Maybe that's a badge of honour in their peer group. Now what could be better for those people than being known and seen on camera for attacking a police officer?

On the other hand: Maybe the story of the data behind these results is another one. As was suggested in the article as well - perhaps the introduction of cameras let violence against police officers to be better measured. Maybe police officers previously did not report all violence against them as doing so would require them to substantiate their claims which is harder if you don't have a video record. With that available reporting went up but not actual incidence of violence.
`r unhide()`

# Exercise 1.2
Guacamole makes you more succesful at online dating?

```{r}
knitr::include_url("https://abcnews.go.com/Lifestyle/adding-extra-guacamole-boost-online-daters-popularity/story?id=51483738")
```


`r hide("Example Answer:")`


The headline suggests that there is a positive causal effect from mentioning Guacamole to receiving responses in online dating. A driver for this is could be that Guacamole is a signal for a healthy lifestyle and health and fitness are qualities that are desired when looking for a romantic partner (The story of the article)

The story of the data could be another:

- Perhaps younger people are more likely to be into guacamole. But younger people can also be expected to get more dating responses. This would imply an upward bias of the Guacamole->Dating success effect.

- A taste for guacamole could be more prevalent among people living in cities. People are more likely to respond to people that are close by. Hence, people in cities will have more people close by and therefore receive more responses. Again this implies an upward bias.




`r unhide()`

