---
title: "Data Stories: Introduction - R examples"
output: word_document
editor_options: 
chunk_output_type: inline

---



# Setup your environment

As I said in class, you have to worry a bit about your working environment. The following chunck shows you 
how you could change that. However, I have simplified things a bit (from what you saw in class) to make it easier for to begin with.

For that to work you have to keep all your data files in a directory called that and all the code files (such as this one) in a directory that sits in the same sub directory. e.g. if you have a directory "data stories course" you want to have in there a directory called "data"" with the data files and a directory "code" with the code files.

Note that by default in a markdown document (i.e. an Rmd document...things are a bit different in a R document) the 
default active working directory is always the directory where the document is stored. So below when I load data,
this is what I assume; e.g. I will load "../data/foreigners.dta". Note that this tells the computer that to find
the file you need to go one level up (that's what ".." means) and then you go down again in the "data" directory
```{r}

print("hello world")
7/6

```

when we print something shows up


```{r setup,include=TRUE}


#mydir=paste0(dropbox_path,'/teaching/data stories')  # set this according to what applies on your computer
mydir=getwd()
print(mydir)
# e.g. mydir="c:/user/Rproject1/"
setwd(mydir)  # This is how you set a working directory in normal R code
getwd()


knitr::opts_knit$set(root.dir = getwd(), verbose = TRUE)  # This makes sure you keep the same directory in subsequent chuncks of code


```
  

# Some playing
```{r , include=TRUE}
  v1=runif(100)
  v2=runif(100)+v1
  
  
  plot(v1,v2)

```


# Migrants
Since they gained power in 2010 the UK Tory Junta have led a relentless campaign of scapegoating immigrants for their policy failures. Part of this is the suggestion that increased levels of immigration lead to more crime.
Here we explore to what extent there is actually any evidence in the data for this.


## Load the Data
```{r }
 getwd()
 
 library(haven)   # make sure libraries such as this are installed. If not go to Tools -> Install Packages

 #df=read_dta("../data/foreigners.dta")
 
#ralf<-read_dta("../data/foreigners.dta")
  
df=read.csv("https://www.dropbox.com/s/g1w75gkw7g91zef/foreigners.csv?dl=1")  

```
some stuff I wrote...
```{r}

print(3/7)

```

## Prepare Data
```{r prepare data , include=TRUE}
  newv=runif(20)
  #df['test']=newv/df$pop11
  summary(df)
  
  summary(df$crimes11)
  summary(df$pop11)
  
  crimes11=df$crimes11
  
  summary(crimes11)

  df['crimesPc']=df$crimes11/df$pop11
  
    
```

## Look at data

A scatter plot
```{r scatter plot , include=TRUE}
plot(df$b_migr11,df$crimesPc)


library(ggplot2)
ggplot(df, aes(x = b_migr11, y = crimesPc)) +
    geom_point()+theme_minimal()+
    xlab("% foreign born")+ylab("# of crimes per capita")


```

How about leaving the City of London out?
```{r no city, include=TRUE}



df[df$crimesPc>5,"area"]


df[df$crimesPc>4.5,"area"]

df[df$crimesPc>4.5,c("area","crimesPc")]


dfx =   df[df$crimesPc<10,]


ggplot(dfx, aes(x = b_migr11, y = crimesPc)) +
    geom_point()+theme_minimal()+
    xlab("% foreign born")+ylab("# of Crimes per capita")

```    
    

What about a correlation?
```{r}
library(dplyr)
#summary(ralf$crimes11)

cor(df[,c("crimes11","crimesPc","b_migr11")],use="complete.obs")

cor(dfx[,c("crimesPc","b_migr11")],use="complete.obs")
df %>% select(crimesPc,b_migr11) %>% cor(use="complete.obs")


library(corrr)
test=df %>% select(crimesPc,b_migr11) 
df %>% select(crimesPc,b_migr11) %>%  correlate() 

```


Or trend line?

```{r scatter plot with line , include=TRUE}
library(ggplot2)
ggplot(dfx, aes(x = b_migr11, y = crimesPc)) +
    geom_smooth(method = "lm", se = FALSE) +
    geom_point()+theme_minimal()+
    xlab("% foreign born")+ylab("# of Crimes per capita")


```








    

```{r regressions}

  
  reg1=lm(crimesPc~b_migr11,df)

  reg1_df=df[-reg1$na.action,]  # get rid of missing values
  
  
  summary(reg1)
  summary(df$b_migr11)
  
  #summary(lm(crimesPc~b_migr11,dfx))
```


Notice that the residulas from a regression are not correlated 

```{r }
  reg1_df['residuals']= reg1$residuals
  

  cor(reg1_df[,c("residuals","b_migr11")],use="complete.obs")


  ggplot(reg1_df, aes(x = b_migr11, y = residuals)) +
    geom_smooth(method = "lm", se = FALSE) +
    geom_point()+theme_minimal()+
    xlab("% foreign born")+ylab("Regression residuals")


  

    
```




# Some further perspectives on the data


Just some basic numbers
```{r}
#library(psych)
#describe(df$crimesPc)
#describe(df$b_migr11)

library(Hmisc)
Hmisc::describe(df[,c("b_migr11","crimesPc")])


summary(df$b_migr11)
summary(df$crimesPc)




library(scales)
library(magrittr)



ggplot(data = df, aes(b_migr11)) + 
  geom_histogram( bins=30,              aes(y = ..density.., fill=..count..)) +
  geom_density()+theme_minimal()


ggplot(data = df, aes(crimesPc)) + 
  geom_histogram( bins=30,              aes(y = ..density.., fill=..count..)) +
  geom_density()+theme_minimal()




```







