---
title: "Multivariate Regression"
output: word_document
editor_options: 
  chunk_output_type: inline
chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




# Wage regression

Let's look at a new example: regressions of hourly wages on years of education (note this is US data)

```{r}
library(haven) 
getwd()  
#load dataset 
data <- read_dta("../data/wage1.dta")
head(data)


summary(data)


mod1 <- lm(wage ~ educ, data)
summary(mod1)
plot(data$educ, data$wage, col = "blue")
lines(data$educ, mod1$fitted.values, col = "red")





summary(lm(exper ~ educ  , data))
summary(lm(wage  ~ educ + exper, data))
summary(lm(educ  ~ female,data))
summary(lm(wage  ~ educ+female,data))
summary(lm(wage  ~ educ+exper+female,data))




```






```{r}
 library(haven)   # make sure libraries such as this are installed. If not go to Tools -> Install Packages
 df=read_dta("../data/foreigners.dta")
 df['crimesPc']=df$crimes11/df$pop11
 
 
 library(dplyr)
 df=df%>%merge(read_dta("../../data/foreigners_more.dta"),by="area") # Getting further variables
 df=df%>%filter(crimesPc<15)
 
 
 reg1=lm(crimesPc~b_migr11,df)
 summary(reg1)
 

 
 
 summary(lm(crimesPc~b_migr11+pop11,df))
 summary(lm(b_migr11~pop11,df))
 
 summary(lm(crimesPc~b_migr11+pop11+urate2011,df))
 summary(lm(b_migr11~urate2011,df))
 
 
 summary(lm(crimesPc~b_migr11+pop11+medianage,df))
 
 
 
 summary(lm(crimesPc~b_migr11+pop11+urate2011+medianage,df))
 
 
``` 
 
 
 
 
 An alternative strategy: Unemployment in 2004 can't be affected by the surge in migration after 2004
```{r}

 summary(lm(crimesPc~b_migr11+pop11+medianage+urate2004,df))
 summary(lm(crimesPc~b_migr11+urate2004+medianage,df))

```
 
 
 
 

```{r}





plot(df$pop11,df$b_migr11)
plot(df$urate2011,df$b_migr11)
plot(df$medianage,df$b_migr11)
#df[df$pop11>600000,"area"]
 
```



How about differentiating by different groups? Muslims' get a lot of hate by English neo nazis
```{r}

reg2=lm(crimesPc~b_migr11+mus_sh+pop11+medianage+urate2011,df)
summary(reg2)

```


What about imperfect colinearity?

```{r}


reg2=lm(crimesPc~b_migr11+pop11+medianage+urate2011+urate2010+urate2012,df)
summary(reg2)


```



  
# Multicolinearity


```{r}
  data["exper_h"] <- data$exper>quantile(data$exper,0.5)  
  head(data)
  data["educ_h"] <- data$educ>quantile(data$educ,0.5)  
  
  table(data$exper_h, data$educ_h)
  
  data2<-subset(data, exper_h==educ_h)
  table(data2$exper_h, data2$educ_h)
  
  summary( lm(wage ~ educ_h + exper_h, data2))
  # # change variable type
  # data$ED <- as.factor(data$ED)
  # 
  # mod2 <- lm(LWAGE ~ ED, data)
  # summary(mod2)

```



What about imperfect colinearity?

```{r}


reg_mc=lm(crimesPc~b_migr11+pop11+medianage+urate2011+urate2010+urate2012,df)
summary(reg_mc)


cor(df[,c("urate2011","urate2010","urate2012")],use="complete.obs")

```

A joint hypothesis test could be a good idea



```{r}
library("car")

linearHypothesis(reg_mc,c("urate2011=0","urate2012=0","urate2010=0"))


```

  
# joint hypothesis tests

```{r}
  wage1 <- read_dta("../data/wage1.dta")
  wage_educ_exper <- lm(wage ~ educ + exper, wage1)
  
  
  #< by transforming model
  
     wage1["educPexper"] <- wage1$educ+wage1$exper
     head(wage1)
  
  #>
  
  summary( lm(wage ~ educPexper + exper, wage1))
  
  
  library("car")
  linearHypothesis(wage_educ_exper,  c("educ =0" ,"exper=0") )
  
```  







