---
title: "Taking back control"
author: Ralf Martin
#output: html_document
output: word_document
editor_options: 
  chunk_output_type: inline
---




# Wage regression

Let's look at a new example: regressions of hourly wages on years of education (note this is US data)

```{r}
library(haven) 
getwd()  
#load dataset 
data <- read.csv("https://www.dropbox.com/s/9agc2vmamfztlel/WAGE1.csv?dl=1")
head(data)


summary(data)


mod1 <- lm(wage ~ educ, data)
summary(mod1)
plot(data$educ, data$wage, col = "blue")
lines(data$educ, mod1$fitted.values, col = "red")





summary(lm(exper ~ educ  , data))
summary(lm(wage  ~ educ + exper, data))
summary(lm(educ  ~ female,data))
summary(lm(wage  ~ educ+female,data))
summary(lm(wage  ~ educ+exper+female,data))




```






```{r}
 library(haven)   # make sure libraries such as this are installed. If not go to Tools -> Install Packages
 df=read.csv("https://www.dropbox.com/s/g1w75gkw7g91zef/foreigners.csv?dl=1")
 df['crimesPc']=df$crimes11/df$pop11
 
 
 library(dplyr)
 df=df%>%merge(read.csv("https://www.dropbox.com/s/gwq2wmmxr8s3v7t/foreigners_more.csv?dl=1"),
               by="area") # Getting further variables


 
 reg1=lm(crimesPc~b_migr11,df)
 
 
 reg1=lm(crimesPc~pop11+b_migr11,df)
 summary(reg1)

 summary(reg1)
 

 
 
 summary(lm(crimesPc~b_migr11+pop11,df))
 summary(lm(b_migr11~pop11,df))
 
 summary(lm(crimesPc~b_migr11+pop11+urate2011,df))
 summary(lm(b_migr11~urate2011,df))
 
 
 summary(lm(crimesPc~b_migr11+pop11+medianage,df))
 
 
 
 summary(lm(crimesPc~b_migr11+pop11+urate2011+medianage,df))
 
 
``` 
 
 
 
 
 An alternative strategy: Unemployment in 2004 can't be affected by the surge in migration after 2004
```{r}

 summary(lm(crimesPc~b_migr11+pop11+medianage+urate2004,df %>% filter(crimesPc<15)))

 summary(lm(crimesPc~b_migr11+urate2011+medianage,df))
 summary(lm(crimesPc~b_migr11+urate2004+medianage,df))

```
 
 
 
 

```{r}





plot(df$pop11,df$b_migr11)
plot(df$urate2011,df$b_migr11)
plot(df$medianage,df$b_migr11)
#df[df$pop11>600000,"area"]
 
```



How about differentiating by different groups? Muslims' get a lot of hate by English neo nazis
```{r}

reg2=lm(crimesPc~b_migr11+mus_sh+pop11+medianage+urate2011,df)
summary(reg2)

```

# Multicolinearity



## Perfect colinearity

```{r}
library(ggplot2)
data=data %>% mutate(educ_in_days=educ*365)
cor(data %>% select(educ,educ_in_days))
ggplot(data,aes(x=educ,y=educ_in_days))+geom_point()

reg2=lm(wage~female+educ+educ_in_days,data)
reg2 %>% summary()


```



## What about imperfect colinearity?



  



```{r}
library(car)
evenmore=read.csv( "https://www.dropbox.com/s/pwotro2ghawkppg/foreign_evenmore.csv?dl=1")
df=df%>% inner_join(evenmore,by="area")



rr=lm(crimesPc~b_migr11+urate2011+
            pop11+
            shxage0t17+ 
            shxage18t29+shxage30t44+shxage45t64+meanage,df  %>% filter(crimesPc<150))
rr%>% summary()
```


```{r}
cor(df %>% select(shxage0t17, 
                  shxage18t29,
                  shxage30t44,
                  shxage45t64,meanage),use="complete.obs")

```



A joint hypothesis test could also be a good idea



```{r}
library("car")

rr%>% vif()

```

  
# Joint hypothesis tests

Testing multiple restrictions at once; e.g. does age really not matter in the regression above?

```{r}

  library("car")
  linearHypothesis(rr,  c("shxage0t17 =0" ,
                          "shxage18t29=0",
                          "shxage30t44=0",
                          "shxage45t64=0",
                          "meanage=0"
                          ) )
```



```{r}
  wage1 <- read.csv("https://www.dropbox.com/s/9agc2vmamfztlel/WAGE1.csv?dl=1")
  wage_educ_exper <- lm(wage ~ educ + exper, wage1)
  
  
  #< by transforming model
  
     wage1["educPexper"] <- wage1$educ+wage1$exper
     head(wage1)
  
  #>
  
  summary( lm(wage ~ educPexper + exper, wage1))
  
  
  library("car")
  linearHypothesis(wage_educ_exper,  c("educ =0" ,"exper=0") )
  
```  







